---
title: "coyote_figures script"
author: "authors: Jamie Clarke, , Marissa Dyck, Larissa Bron, Madison Carlson, Sophia Labiy, Zoe Penno, Hayley Webster, Jason T Fisher"
date: "last updated: 30/12/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## about
This is walk-through on how to make figures and maps for *the coyote manuscript*. It was written by Jamie Clarke with help from Dr. Marissa Dyck. Before running this script, you will need to format and analyze the data using the coyote_formatting and coyote_analysis scripts. The spatial data required to run the script is available on Jamie's GitHub at https://github.com/jfclarke/osm_coyote_glmm, or in the links provided. Happy figure- and map-making :-)

## 1) set-up

#### start by loading in relevant packages:
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(PerformanceAnalytics)
library(ggplot2)
library(cowplot)
library(sf)
library(dplyr)
library(tmap)
library(tmaptools)
library(ggpubr)
library(grid)
library(sp)
library(lme4)
library(ggeffects)
library(scales)
library(rphylopic)
library(MuMIn)
```

#### and setting the ggplot theme
```{r ggplot_theme, message=FALSE}
theme_set(theme_classic())
```

## 2) data import

#### a) read in processed data (created using coyote_formatting script):
```{r coyote_data, message=FALSE}
coyote_data <-
  
  read_csv('data/processed/coyote_data.csv') %>% 
  
  # add wide_linear column combining all wide linear features (determined using coyote_analysis script)
  mutate(wide_linear = 
           roads + 
           seismic_lines +
           transmission_lines)
```

#### b) run top model (determined using coyote_analysis script):
```{r global, message=FALSE}
global <-
  
  glmer(
    cbind(coyote_pres, coyote_abs) ~
      scale(nat_land) +
      scale(wide_linear) +
      scale(white_tailed_deer) +
      scale(moose) +
      scale(red_squirrel) +
      scale(snowshoe_hare) +
      scale(grey_wolf) +
      scale(lynx) +
      scale(fisher) +
      (1 | array),
    data = coyote_data,
    family = binomial)
```

## 3) calculating odds ratios

```{r global_odds, message=FALSE}
global_odds <-
  
  # calculate confidence intervals
  confint(global,
          parm = 'beta_') %>% 
  
  # extract fixed effects coefficients
  cbind(est = fixef(global)) %>% 
  
  # exponentiate to get odds ratios
  exp() %>% 
  
  as.data.frame() %>% 
  
  # name first column and preserve it for next step
  rownames_to_column(var = 'term') %>% 
  
  # convert to tibble for easier manipulation
  as_tibble() %>%
  
  # remove intercept information
  filter(term != '(Intercept)') %>% 
  
  # rename confidence %s
  rename(lower = '2.5 %',
         upper = '97.5 %') %>% 
  
  # add a column with cleaned-up plot label names (so we don't have to do this in ggplot2)
  add_column(label = c('natural landcover',
                       'wide linear features',
                       'white-tailed deer',
                       'moose',
                       'red squirrel',
                       'snowshoe hare',
                       'grey wolf',
                       'lynx',
                       'fisher')) %>% 
  
  # change label column into a factor for plotting
  mutate(label = as.factor(label)) %>% 
  
  # add column for whether error bars overlap with 1 (for plotting)
  add_column(overlap = c('N',
                         'N',
                         'Y',
                         'Y',
                         'Y',
                         'N',
                         'N',
                         'N',
                         'Y'))
```

## 4) odds ratios plot
```{r odds_plot}
odds_plot <-
  
  ggplot(data = global_odds %>%
           
           # customize order of labels
           mutate(label = fct_relevel(label,
                                      'natural landcover',
                                      'wide linear features',
                                      'fisher',
                                      'lynx',
                                      'grey wolf',
                                      'red squirrel',
                                      'snowshoe hare',
                                      'white-tailed deer',
                                      'moose')),
         aes(x = label,
             y = est,
             color = overlap)) +

  # add background for landscape covariates
  annotate('rect',
           xmin = 7.5,
           xmax = 10,
           ymin = -Inf,
           ymax = Inf,
           fill = 'darkseagreen',
           alpha = 0.4) +
  
  # add background for competitor covariates
  annotate('rect',
           xmin = 4.5,
           xmax = 7.5,
           ymin = -Inf,
           ymax = Inf,
           fill = 'lightsteelblue1',
           alpha = 0.4) +
  
  # add background for prey covariates
  annotate('rect',
           xmin = 0,
           xmax = 4.5,
           ymin = -Inf,
           ymax = Inf,
           fill = 'tomato',
           alpha = 0.2) +
  
  # add line at 1
  geom_hline(yintercept = 1,
             alpha = 0.7) +
  
  # add points for estimates
  geom_point(size = 3) +
  
  # add confidence intervals
  geom_linerange(aes(ymin = lower,
                     ymax = upper),
                 linewidth = 0.5) +
  
  # set colour based on whether error bars overlap with 1
  scale_color_manual(values = c('black', 'grey70' ),
                     labels = c('N', 'Y')) +
  
  # reverse the order of labels on the x axis (ggplot annoyingly plots them in reverse order)
  scale_x_discrete(limits = rev) +
  
  # rename y axis title
  ylab('odds ratio') +
  
  # flip x and y axis 
  coord_flip() +
  
  # specify theme elements
  theme(panel.grid = element_blank(),
        axis.title.y = element_blank(),
        legend.position = 'none',
        axis.title.x = element_text(size = 14))

# run odds_plot to see figure
odds_plot
```

## 5) export odds ratio plot

#### save odds_plot to 'figures' folder:
```{r odds_plot_exp, message=FALSE}
ggsave('odds_ratio_plot.tiff',
       odds_plot,
       path = 'figures')
```

## 6) determining predicted probabilities

#### predicted probabilities given each covariates of interest:
```{r pp, message=FALSE}
pp_nat_land <-
  
  ggpredict(global,
            terms = 'nat_land [all]',
            type = 'fe') # fe = fixed effects

pp_wide_linear <-
  
  ggpredict(global,
            terms = 'wide_linear [all]',
            type = 'fe')

pp_moose <-
  
  ggpredict(global,
            terms = 'moose [all]',
            type = 'fe')

pp_red_squirrel <-
  
  ggpredict(global,
            terms = 'red_squirrel [all]',
            type = 'fe')

pp_snowshoe_hare <-
  
  ggpredict(global,
            terms = 'snowshoe_hare [all]',
            type = 'fe')

pp_white_tailed_deer <-
  
  ggpredict(global,
            terms = 'white_tailed_deer [all]',
            type = 'fe')

pp_fisher <-
  
  ggpredict(global,
            terms = 'fisher [all]',
            type = 'fe')

pp_grey_wolf <-
  
  ggpredict(global,
            terms = 'grey_wolf [all]',
            type = 'fe')

pp_lynx <-
  
  ggpredict(global,
            terms = 'lynx [all]',
            type = 'fe')
```

## 7) predicted probabilities plots

#### <span style="color: darkblue;">side note:</span> to find the PhyloPic uuid (= image ID), type the following code in the console:
```{r phylopic, eval=FALSE}
pick_phylopic(name = 'Alces alces', n = 2) # or whatever other species/number
```
#### where name is the scientific or common name of the species of interest and n is the number of options you want to peruse; when you find the image you want, select it and find the uuid in the console output

plot predicted probability of coyote occurrence given proportion of natural landcover:
```{r plot_nat_land}
plot_nat_land <- 
  
  ggplot(pp_nat_land,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'darkseagreen',
              alpha = 0.4) +
  
  scale_x_continuous(limits = c(0.7, 1),
                     breaks = seq(0.7, 1, by = 0.1),
                     expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 0.65, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('proportion natural land') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35))
```
  
plot predicted probability of coyote occurrence given proportion of wide linear features:
```{r plot_wide_linear}
plot_wide_linear <- 
  
  ggplot(pp_wide_linear,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'darkseagreen',
              alpha = 0.4) +
  
  scale_x_continuous(expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 0.65, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('proportion of wide LFs') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35))
```

plot predicted probability of coyote occurrence given total independent moose detections:
```{r plot_moose}
plot_moose <-
  
  ggplot(pp_moose,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'tomato',
              alpha = 0.4) +
  
  scale_x_continuous(expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total moose detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add moose silhouette
  add_phylopic(uuid = '74eab34a-498c-4614-aece-f02361874f79',
               x = 10,
               y = 0.55,
               height = 0.2)
```

plot predicted probability of coyote occurrence given total independent red squirrel detections:
```{r plot_red_squirrel}
plot_red_squirrel <-
  
  ggplot(pp_red_squirrel,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'tomato',
              alpha = 0.4) +
  
  scale_x_continuous(limits = c(0, 150),
                     breaks = seq(0, 150, by = 50),
                     expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total squirrel detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add squirrel silhouette
  add_phylopic(uuid = 'dad08fea-5263-4f57-a37b-c27cbe0eb9a5',
               x = 30,
               y = 0.54,
               height = 0.16)
```

plot predicted probability of coyote occurrence given total independent snowshoe hare detections:
```{r plot_snowshoe_hare}
plot_snowshoe_hare <-
  
  ggplot(pp_snowshoe_hare,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'tomato',
              alpha = 0.4) +
  
  scale_x_continuous(expand = c(0,0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total hare detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add hare silhouette
  add_phylopic(uuid = '44f9db61-88e2-4a44-82f5-57dea9b3798d',
               x = 50,
               y = 0.55,
               height = 0.2)
```

plot predicted probability of coyote occurrence given total independent white-tailed deer detections:
```{r plot_white_tailed_deer}
plot_white_tailed_deer <-
  
  ggplot(pp_white_tailed_deer,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'tomato',
              alpha = 0.4) +
  
  scale_x_continuous(limits = c(0, 130),
                     breaks = seq(0, 130, by = 30),
                     expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total deer detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add deer silhouette
  add_phylopic(uuid = '4584be20-4514-4673-a3e8-97e2a6a10e57',
               x = 29,
               y = 0.53,
               height = 0.24)
```

plot predicted probability of coyote occurrence given total independent fisher detections:
```{r plot_fisher}
plot_fisher <-
  
  ggplot(pp_fisher,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'lightsteelblue1',
              alpha = 0.4) +
  
  scale_x_continuous(limits = c(0, 12.5),
                     breaks = seq(0, 12.5, by = 3),
                     expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total fisher detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add fisher silhouette
  add_phylopic(uuid = '735066c6-2f3e-4f97-acb1-06f55ae075c9',
               x = 3.5,
               y = 0.55,
               height = 0.12)
```

plot predicted probability of coyote occurrence given total independent grey wolf detections:
```{r plot_grey_wolf}
plot_grey_wolf <-
  
  ggplot(pp_grey_wolf,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'lightsteelblue1',
              alpha = 0.4) +
  
  scale_x_continuous(limits = c(0, 13),
                     breaks = seq(0, 13, by = 3),
                     expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total wolf detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add wolf silhouette
  add_phylopic(uuid = '8cad2b22-30d3-4cbd-86a3-a6d2d004b201',
               x = 3.5,
               y = 0.55,
               height = 0.17)
```

plot predicted probability of coyote occurrence given total independent lynx detections:
```{r plot_lynx}
plot_lynx <-
  
  ggplot(pp_lynx,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes()) +
  
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = 'lightsteelblue1',
              alpha = 0.4) +
  
  scale_x_continuous(expand = c(0, 0)) +
  
  scale_y_continuous(limits = c(0, 0.65),
                     breaks = seq(0, 1, by = 0.2),
                     expand = c(0, 0)) +
  
  xlab('total lynx detections') +
  
  ylab(' ') +
  
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 7),
        axis.title.x = element_text(size = 9),
        axis.title.y.left = element_text(size = 35)) +
  
  # add lynx silhouette
  add_phylopic(uuid = '27a2173a-5903-46fc-83c5-29ed7f421046',
               x = 5.9,
               y = 0.55,
               height = 0.17)
```

## 8) predicted probabilities plot panel

#### combine predicted probabilities plots together in one figure:
```{r pp_plot, warning=FALSE}
pp_plot <-
  
  ggarrange(plot_nat_land,
            plot_wide_linear,
            plot_fisher,
            plot_lynx,
            plot_grey_wolf,
            plot_red_squirrel,
            plot_snowshoe_hare,
            plot_white_tailed_deer,
            plot_moose,
            labels = 'auto',
            label.x = 0.88,
            font.label = list(size = 12),
            ncol = 5,
            nrow = 2) %>% 
  
  annotate_figure(left = text_grob('predicted coyote occurrence',
                                   rot = 90,
                                   vjust = 0.5),
                  fig.lab.size = 14)

# run pp_plot to see figure
pp_plot
```

## 9) export predicted probabilities figure

#### save pp_plot to 'figures' folder:
```{r pp_plot_exp, message=FALSE}
ggsave('predicted_probabilities_panel.png',
       pp_plot,
       path = 'figures',
       width = 250,
       height = 100,
       units = 'mm',
       bg = 'white')
```

## 10) predicted probabilities w/ random effects

#### determining and plotting an example graph of predicted probabilities with random effects
  
#### choosing wide linear features as the fixed effect:
```{r re_pp_wide_linear}
re_pp_wide_linear <-
  
  ggpredict(global, 
            terms = c('wide_linear [all]', # fixed effect
                      'array'), # random effect
            type = "random")
```

#### plot predicted probability of coyote occurrence given total independent lynx detections and random effects of array:
```{r re_plot}
re_plot <-
  
  ggplot(re_pp_wide_linear,
         aes(x = x,
             y = predicted)) +
  
  geom_line(aes(color = group)) +
  
  # using colourblind-friendly Tol colour scheme
  # more info at https://personal.sron.nl/~pault/
  scale_fill_manual(values = c('#4477AA',
                               '#66CCEE',
                               '#228833',
                               '#CCBB44',
                               '#EE6677',
                               '#AA3377')) +
  
  geom_ribbon(aes(fill = group,
                  ymin = conf.low,
                  ymax = conf.high),
              alpha = 0.1) +
  
  scale_color_manual(values = c('#4477AA',
                                '#66CCEE',
                                '#228833',
                                '#CCBB44',
                                '#EE6677',
                                '#AA3377')) +
  
  labs(color = NULL,
       fill = NULL,
       x = 'proportion of wide LFs',
       y = 'predicted coyote occurrence')

# run re_plot to see figure
re_plot
```

#### save plot:
```{r re_plot_exp, message=FALSE}
ggsave('re_predicted_probability.png',
       re_plot,
       path = 'figures',
       width = 183,
       height = 100,
       units = 'mm')
```

## 11) variance inflation factor

#### calculate variance inflation factors (VIFs):
```{r vif}
vif_model <-
  
  # determine VIFs
  car::vif(global) %>% 
  
  # convert into a format that can be plotted
  as_tibble() %>% 
  
  # add a column labelling covariate name
  add_column(label = c('natural landcover',
                       'wide linear features',
                       'fisher',
                       'lynx',
                       'grey wolf',
                       'red squirrel',
                       'snowshoe hare',
                       'white-tailed deer',
                       'moose')) %>% 
  
  # add a column specifying covariate type
  add_column(cov_type = c('land',
                          'land',
                          'comp',
                          'comp',
                          'comp',
                          'prey',
                          'prey',
                          'prey',
                          'prey'))
```

#### plot VIFs:
```{r vif_plot}
vif_plot <-
  
  ggplot(data = vif_model %>% 
           
           # customize order of labels
           mutate(label = fct_relevel(label,
                                      'natural landcover',
                                      'wide linear features',
                                      'fisher',
                                      'lynx',
                                      'grey wolf',
                                      'red squirrel',
                                      'snowshoe hare',
                                      'white-tailed deer',
                                      'moose')),
         aes(x = label,
             y = value,
             fill = cov_type)) +
  
  geom_col() +
  
  # set colour based on covariate type
  scale_fill_manual(values = c('lightsteelblue1', 'darkseagreen', 'tomato'),
                     labels = c('land', 'comp', 'prey')) +
  
  # add line at 1
  geom_hline(yintercept = 1,
             alpha = 0.5,
             linetype = 'dashed') +
  
  # reverse the order of labels on the x axis (ggplot annoyingly plots them in reverse order)
  scale_x_discrete(limits = rev) +
  
  scale_y_continuous(limits = c(0, 1.35),
                     breaks = seq(0, 1.3, by = 0.5),
                     expand = c(0, 0)) +
  
  # rename x axis title
  ylab('VIF values') +
  
  # flip x and y axis 
  coord_flip() +
  
  # specify theme elements
  theme(axis.title.y = element_blank(),
        legend.position = 'none',
        axis.title.x = element_text(size = 14))

# run vif_plot to see figure
vif_plot
```

#### save plot:
```{r vif_plot_exp}
ggsave('vif_plot.png',
       vif_plot,
       path = 'figures',
       width = 100,
       height = 100,
       units = 'mm')
```


## 12) mapping set-up

#### a) read in LU polygons where CTs were deployed:
```{r lus, message=FALSE}
# first: specify which LUs
# for 2021-2022 - LUs 2 + 3
# for 2022-2023 - LUs 1, 12, 15 + 21
target <- c(2, 3, 1, 13, 15, 21)

# then: read in data + filter for those LUs
lus <-
  
  st_read('data/spatial/ALL_LUs.shp') %>% 
  
  set_names(
    names(.) %>% 
      tolower()) %>% 
  
  st_transform(crs = 26912) %>% 
  
  filter(lu %in% target)
```

#### b) read in CT deployment data for 2022:
```{r cts_2022, message=FALSE}
cts_2022 <-
  
  st_read('data/spatial/OSM2022_SITES_ALL.shp') %>% 
  
  set_names(
    names(.) %>% 
      tolower())
```

#### c) read in CT deployment data for 2021 per LU and combine into single dataframe:
```{r cts_2021, message=FALSE}
cts_2021 <-
  
  st_read('data/spatial/OSM_2021_LU2.shp') %>%
  rbind(st_read('data/spatial/OSM_2021_LU3.shp')) %>%
  
  set_names(
    names(.) %>% 
      tolower())
```

#### d) read in provincial boundaries:
```{r provs, message=FALSE}
# source = Statistics Canada (https://www12.statcan.gc.ca/census-recensement/2011/geo/bound-limit/bound-limit-2016-eng.cfm)
provs <- 
  
  st_read('data/spatial/lpr_000b16a_e.shp') %>% 
  
  set_names(
    names(.) %>% 
      tolower()) %>% 
  
  st_transform(crs = 26912)

```

#### e) subset Alberta:
```{r ab_boundary}
ab_boundary <-
  
  provs %>% 
  filter(prename == 'Alberta')
```

#### f) crop Alberta spatial info to northeast (oil sands region):
```{r ne_ab}
# figure out spatial extent of LUs
raster::extent(lus)

# make bounding box of NE Alberta based on extents
ne_ab_bbox <- 
  
  st_bbox(c(xmin = 365982.3,
            xmax = 563572,
            ymin = 6036147,
            ymax = 6356412),
          crs = st_crs(26912))
```

#### g) read in Alberta roads layer:
```{r roads, message=FALSE}
# source = National Road Network (https://open.canada.ca/data/en/dataset/cb4911f1-89d8-47f0-92e1-2cbfac3d300b/resource/e7fa72ef-92ea-4465-bd8f-9b6b765ae0ec)
roads <- 
  
  st_read('data/spatial/NRN_AB_14_0_ROADSEG.shp') %>% 
  
  set_names(
    names(.) %>% 
      tolower()) %>% 
  
  st_transform(crs = 26912) %>% 
  
  st_union() # simplify layer to speed up mapping
```

#### h) read in cities layer:
```{r cities, message=FALSE}
cities <-
  
  st_read('data/spatial/city_points.shp') %>% 
  
  set_names(
    names(.) %>% 
      tolower()) %>% 
  
  st_transform(crs = 26912) %>% 
  
  filter(name == 'Fort McMurray')
```
 
## 13) map of study area

#### extend bounding box so map features (e.g., scale bar) don't overlap with map itself:
```{r bbox1}
bbox1 <-  st_bbox(ne_ab_bbox)

xrange <- bbox1$xmax - bbox1$xmin

bbox1[1] <- bbox1[1] - (0.4 * xrange) # xmin - left

bbox1 <- bbox1 %>%
  st_as_sfc() # make bounding box an sf polygon
```

#### set tmap mode to 'plot':
```{r tmap_mode, message=FALSE}
tmap_mode('plot')
```

#### map LUs and CT deployments in NE Alberta:
```{r lus_cts}
lus_cts <-
  
  # this establishes the spatial extent of the map
  tm_shape(ab_boundary,
           bbox = bbox1) +
  tm_fill() +
  
  # map neighbouring provinces in the background
  tm_shape(provs) +
  tm_fill(col = 'grey20') +
  
  # layer Alberta on top
  tm_shape(ab_boundary,
           bbox = ne_ab_bbox) +
  tm_fill(col = 'grey70') +
  
  # add a polygon for LU1
  tm_shape(lus %>% 
             filter(namerefere == 'LU1')) +
  tm_fill(col = 'grey95') +
  
  # add a polygon for LU2
  tm_shape(lus %>% 
             filter(namerefere == 'LU2')) +
  tm_fill(col = 'grey95') +
  
  # add a polygon for LU3
  tm_shape(lus %>% 
             filter(namerefere == 'LU3')) +
  tm_fill(col = 'grey95') +
  
  # add a polygon for LU13
  tm_shape(lus %>% 
             filter(namerefere == 'LU13')) +
  tm_fill(col = 'grey95') +
  
  # add a polygon for LU15
  tm_shape(lus %>% 
             filter(namerefere == 'LU15')) +
  tm_fill(col = 'grey95') +
  
  # add a polygon for LU21
  tm_shape(lus %>% 
             filter(namerefere == 'LU21')) +
  tm_fill(col = 'grey95') +
  
  # layer roads and cities on top
  tm_shape(roads) +
  tm_lines(col = 'grey40',
           lwd = 0.5) +
  
  tm_shape(cities) + 
  tm_dots(col = 'black',
          size = 0.3) +
  tm_text('name',
          size = 0.7,
          just = c(-0.11,-0.11)) +
  
  # add points for CT deployment locations in LU1
  tm_shape(cts_2022 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU1'))) +
  tm_dots(col = '#005AB5',
          size = 0.02) +
  
  # add points for CT deployment locations in LU2
  tm_shape(cts_2021 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU2'))) +
  tm_dots(col = '#DC3220',
          size = 0.02) +
  
  # add points for CT deployment locations in LU3
  tm_shape(cts_2021 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU3'))) +
  tm_dots(col = '#DC3220',
          size = 0.02) +
  
  # add points for CT deployment locations in LU13
  tm_shape(cts_2022 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU13'))) +
  tm_dots(col = '#005AB5',
          size = 0.02) +
  
  # add points for CT deployment locations in LU15
  tm_shape(cts_2022 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU15'))) +
  tm_dots(col = '#005AB5',
          size = 0.02) +
  
  # add points for CT deployment locations in LU21
  tm_shape(cts_2022 %>% 
             crop_shape(lus %>% 
                          filter(namerefere == 'LU21'))) +
  tm_dots(col = '#005AB5',
          size = 0.02) +
  
  # add scale bar
  tm_scale_bar(position = c('left', 'BOTTOM'),
               text.size = 0.7,
               width = 0.2) +
  
  # specify layout elements
  tm_layout(frame = FALSE,
            bg.color = 'transparent')
```

#### set NE Alberta bbox as an sf object:
```{r ne_ab_bbox}
ne_ab_bbox <- 
  ne_ab_bbox %>% 
  st_as_sfc()
```

#### overlay map of Alberta within Canada, with box around NE Alberta:
```{r ne_ab_can}
ne_ab_can <-
  tm_shape(provs) +
  tm_fill(col = 'grey20') +
  
  tm_shape(ab_boundary) +
  tm_fill(col = 'grey70') +
  
  tm_shape(ne_ab_bbox) +
  tm_borders(col = 'black',
             lwd = 1) +
  
  tm_layout(frame.lwd = 2)
```

#### overlay maps:
```{r inset_map, warning=FALSE}
# convert tmap objects into elements that cowplot can recognize
lus_cts_grob <- tmap_grob(lus_cts)
ne_ab_can_grob <- tmap_grob(ne_ab_can)

# now: inset using cowplot
inset_map <-
  
  ggdraw(lus_cts_grob) + # add first layer = study area map
  draw_plot(ne_ab_can_grob, # inset map of Canada and specify position
            x = 0.13,
            y = 0.69,
            width = 0.27,
            height = 0.27) +
  theme(panel.background =
          element_rect(
            fill = 'transparent'),
        panel.border = 
          element_rect(
            colour = NA,
            fill = NA))

# run inset_map to see figure
inset_map
```

## export study area map with inset

#### save inset_map to 'figures' folder:
```{r inset_map_exp, message=FALSE}
ggsave('study_area_map.png',
       inset_map,
       path = 'figures',
       width = 100,
       height = 105,
       units = 'mm',
       bg = 'transparent')
```

